\documentclass[a4paper]{article}

\usepackage{amsmath, graphicx, float, blindtext} % for dummy text
\graphicspath{ {./images/} }
\title{Suggested Readings}
\author{Shubham Gupta}

\begin{document}
\maketitle
\section{Incrementality in Deterministic Dependency Parsing}
\begin{itemize}
    \item Why incrementality?
        \begin{itemize}
            \item Practical and can be used in real time applications
            \item Theoritical i.e human parsing is largely incremental, so it makes sense to try it out in the text as well.
            \item \textbf{Core aim}: Build a complete syntactic analysis for the input string
            \item Dependencies can be done by labelling each word in the sentence by POS tags and the arcs between each of these words with their gramatical function
            \item Parser configs represented by triples (S, I, A). S = Stack, I = List of input tokets, A = current arc relation in dependency graph
            \item \textbf{Incremental}: For bottom up parsers, idea is to have a term related to graph at all times. Number of elements in stack is not greater than 2 
            \item \textbf{Arc eager parsing}: Do both bottom-up and top-down parsing. Left dependents bottom-up and right dependents top-down 
            \item In standard algorithm, Right-redice is bottom-up approach i.e it is a Right-Arc + Reduce function.
            \item \textbf{Incremental for arc parsing}: Enforce that the graph should be connected at all times, since two tokens need not be present on the stack to create the dependency  
            \item \textit{Top Down parsing}: Head is attached to dependant before dependent is attached to its other dependents 
            \item \textit{Bottom up parsing}: Dependent is attached to its head before head is attached to its parent's head.
            \item \textbf{Performance}: Evaluation based on arc-eager parsing with memory-based classifier for predicting next transition
            \item 90\% of all confihs ave no more than three connected components on the stack.
            \item Sentences that violate the incremental parsing property are because:
                \begin{itemize}
                    \item Can't be parse into a well formed tree i.e single projective dependency tree
                \end{itemize}
        \end{itemize}
\end{itemize}
\section{Conclusion}
\begin{itemize}
    \item This paper shows that strict incremental parsing is not acheivble with the framework presented.
    \item HOWEVER, arc-eager parsing algorithm is optimial for incremental parsing.
\end{itemize}
\end{document}
